---
title: "Analysis of Netflix Movies and TV Shows"
author: "Nicole Pasternak"
format:
  revealjs:
    theme: simple
    scrollable: true
    center: false
    slide-level: 2
execute:
  echo: true
---

## Introduction

This project analyzes a dataset of Netflix movies and TV shows containing over 8,000 titles and variables such as content type, release year, and rating. The goal of this analysis is to apply descriptive statistics and data visualization techniques to identify patterns in Netflix’s content library.

In particular, this analysis focuses on:

-   Content type (movies vs. TV shows)
-   Distribution of release years
-   Distribution of content ratings
-   Dataset source: Kaggle Netflix Titles dataset (<https://www.kaggle.com/datasets/shivamb/netflix-shows>)

By examining these variables, we aim to better understand how Netflix’s catalog is structured and how its content has evolved over time. This project demonstrates how R can be used to import, summarize, and visualize real-world data to uncover meaningful patterns.

## Load Data

```{r}
library(tidyverse)
netflix <- read.csv("netflix_titles.csv")
head(netflix)
dim(netflix)
```

This step loads the necessary R package for data analysis and imports the Netflix dataset into R as a data frame called netflix. The output displays the first six rows of the dataset and its dimensions, allowing us to confirm that the data has been correctly loaded and structured for analysis.

## Data Description

Each row in the dataset represents a single Netflix title (either a movie or a TV show).

The variables used in this analysis can be classified as follows:

**Categorical variables:**

-   Type: Indicates whether a title is a Movie or TV Show\
-   Rating: Content rating classification (e.g., TV-MA, TV-14)

**Quantitative variable:**

-   "release_year": Year the title was originally released

## Summary Statistics (Release Year)

```{r}
mean(netflix$release_year, na.rm = TRUE)
median(netflix$release_year, na.rm = TRUE)
summary(netflix$release_year)
```

The summary statistics show that release years are heavily concentrated in recent decades. The median release year is more informative than the mean because the distribution is skewed toward newer content, indicating that most titles in the dataset were released in the modern streaming era.

## Distribution of Content Type

```{r}
#| fig-width: 12
#| fig-height: 7
#| out-width: "100%"
#| dpi: 150

ggplot(netflix, aes(x = type)) +
  geom_bar() +
  labs(
    title = "Distribution of Netflix Content Types",
    x = "Type",
    y = "Count"
  )
```

## Understanding

The dataset contains more movies than TV shows, indicating that movies make up the majority of [Netflix]{.underline}’s catalog in this sample.

## Distribution of Release Year

```{r}
#| fig-width: 7
#| fig-height: 4
#| out-width: "100%"

ggplot(netflix, aes(x = release_year)) +
  geom_histogram(binwidth = 5) +
  labs(
    title = "Distribution of Release Years",
    x = "Release Year",
    y = "Count"
  )
```

## Examination

The distribution shows a clear right-skew toward recent years, suggesting strong growth in content production in the streaming era.

## Distribution of Ratings

```{r}
#| fig-width: 7
#| fig-height: 3.8
#| out-width: "100%"

ggplot(netflix, aes(x = rating)) +
  geom_bar() +
  labs(
    title = "Distribution of Content Ratings",
    x = "Rating",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Analysis

The rating distribution is uneven, with TV-MA and TV-14 appearing most frequently. This indicates that the content in the dataset is primarily targeted toward mature and teen audiences.

## External Image

![Netflix Logo](https://upload.wikimedia.org/wikipedia/commons/0/08/Netflix_2015_logo.svg)

-   This image provides visual context for the dataset by representing Netflix, the platform from which the data originates. While it does not contain numerical information, it helps connect the analysis to its real-world source.

## Key Findings

Overall, the results consistently show that the dataset is dominated by movies, concentrated in recent years, and primarily composed of mature-rated content.
